### 2017-03-12 CO750-ApproximationAlgorithmsTopics/assignments/assignment_02.tex
3a

>>>>

%Q3
Consider again the problem $(IP)$, associated with a graph $G=(V,E)$ with edges costs $c: E \rightarrow \R_+$ given by:
x&\in \Z^{|E|}.
Suppose that $f:2^V \rightarrow \N$ is symmetric, meaning for all $S\subseteq V$:
$$f(S) = f(V\backslash S)$$
and that $f$ is subadditive, meaning for all $A,B \subseteq V$:
$$f(A\cup B) \leq f(A) + f(B).$$
We also suppose that the graph $G$ is a cycle on $n$ vertices.
%3a
\subsection*{a}
We begin by working towards a compact formulation for $(IP)$ on $G$. Since $G$ is a cycle we may label the vertices of $G$ as $V(G) = \{v_1,\dots, v_n\}$ such that $v_iv_j \in E(G)$ if and only if $$j \equiv i\pm1\mod n$$. For $i, \ell \in [n]$ ($[n] := \{1,\dots, n\}$) we define the path from $i$ of length $\ell$ as 
$$P(i, \ell) := \{v_i, v_{i+1}, \dots, v_{i + \ell} \}$$
where addition is taken modulo $n$. We use the convention that when $\ell = n$ we have the entire cycle, and when $\ell = 0$ we have just $v_i$. We define the family of all paths on $G$ as
$$\cP := \{P(i,\ell) : i, \ell \in [n]\}.$$
Observe that the size of $\cP$ is polynomial in $n$, in particular $|\cP| = O(n^2)$ (since $n$ choices for $i$ and $n$ choices for $\ell$ covers all possible $P(i,\ell)$ with some overlap).
\begin{lemma}\label{lemma:compact}
The problem $(IP)$ on $G$ is equivalent to the compact integer program $(\overline{IP})$ given by:
\text{s.t.}\  \sum_{e\in\delta(P)} x_e &\geq f(P) &\forall P \in \cP \\
x&\in \Z^{|E|}.
Since the objective functions are the same, we need only demonstrate that the feasible regions of $(IP)$ and $(\overline{IP})$ are equal. Since each $P \in \cP$ is a subset of $V$, it is clear that the feasible region of $(IP)$ is a subset of the feasible region of $(\overline{IP})$. So we just have to show the reverse containment.
Let $x$ be a feasible solution for $(\overline{IP})$. Let $S \subseteq V$. If we consider the connected components of $G[S]$ we observe that they are all paths. Hence there exist disjoint $P_1, \dots, P_k \in \cP$ such that
$$S = P_1 \dot\cup \dots \dot\cup P_k$$
where each $P_i$ corresponds to a connected component of $G[S]$. Now we see that
f(S) &= f(P_1 \dot\cup \dots \dot \cup P_k) \\
&\leq f(P_1) + \dots + f(P_k) &\text{by subadditivity} \\
&\leq x(\delta(P_1)) + \dots x(\delta(P_k)) &\text{by feasibility of $x$}.
We claim that the edges incident to each path are disjoint. To see this, suppose that we have $P_i$ and $P_j$ with $\delta(P_i) \cap \delta(P_j) \neq \emptyset$. Let $v_iv_j \in \delta(P_i) \cap \delta(P_j)$. Since $P_i$ and $P_j$ are disjoint we may assume $v_i \in P_i \backslash P_j$ and $v_j \in P_j \backslash P_i$. But then $P_i$ and $P_j$ are connected by $v_iv_j$ violating that they are distinct connected components. So the claim holds. Thus we observe
f(S) &\leq x(\delta(P_1)) + \dots x(\delta(P_k)) \\
&= x(\bigcup_{i=1}^k \delta(P_i)) &\text{by disjointness}\\
&= x(\delta(S)) &\text{by disjointness}.
Therefore $x$ satisfies the cut constraint for all $S$, and hence $x$ is feasible for $(IP)$.
Hence by lemma \ref{lemma:compact} it is equivalent for us to solve $(\overline{IP})$. Now we define a linear programming relaxation for $(\overline{IP})$, which we will denote $(\overline{LP})$:
\text{s.t.}\  \sum_{e\in\delta(P)} x_e &\geq f(P) &\forall P \in \cP \\
Since $(\overline{LP})$ has a polynomial number of constraints and variables it can be solved in polynomial time. We now define our rounding-based approximation algorithm, $\cA$, for $(\overline{IP})$:
\item Solve $(\overline{LP})$ on $G$  returning solution $\bar{x}$.
\item Let $\bar{x}_I \in \Z^{|E|}$ denote the integral part of $\bar{x}$. Let $\bar{x}_F \in [0,1)^{|E|}$ denote the fractional part of $\bar{x}$. Thus $\bar{x} = \bar{x}_I + \bar{x}_F.$
\item Construct solution $x \in \Z^{|E|}$ by rounding. For each $e \in E$ set:
$$x_e = \begin{cases}
\bar{x}_I + 1, &\text{if } x_F \geq 0.5 \\
\bar{x}_I, &\text{otherwise.}
\item Return solution $x$.
\begin{lemma}\label{lemma:feasible-rounding}
If $\bar{x}$ is a feasible solution to $(\overline{LP})$ then the $x$ obtained from rounding $\bar{x}$ in step $3$ of $\cA$ is a feasible solution to $(\overline{IP})$.
It is immediate from the description of step $3$ that $x$ is integral and non-negative. We need only verify the cut constraints. Let $P \in \cP$. Let $v_i, v_j \in P$ be the start and end vertices of $P$. That is,
$$\delta(P) = \{v_{i-1}v_i,  v_jv_{j+1}\}.$$
For convenience let $e_i = v_{i-1}v_i$ and let $e_j = v_jv_{j+1}$. Since $\bar{x}$ is feasible for $(\overline{LP})$:
$$ f(P)\leq \bar{x}(e_i) + \bar{x}(e_j) = \bar{x}_I(e_i) + \bar{x}_F(e_i)  + \bar{x}_I(e_j) + \bar{x}_F(e_j).$$
Since $f$ maps to $\N$ we have:
$$f(P) \leq \floor{\bar{x}_I(e_i) + \bar{x}_F(e_i)  + \bar{x}_I(e_j) + \bar{x}_F(e_j)}.$$
We consider the following cases: Case 1: $\bar{x}_F(e_i) + \bar{x}_F(e_j) < 1$ and Case 2: $\bar{x}_F(e_i) + \bar{x}_F(e_j) \geq 1$.
In Case 1 the fractional parts vanish during the floor-ing operation and hence
$$ f(P) \leq \bar{x}_I(e_i) + \bar{x}_I(e_j)  \leq x(e_i) + x(e_j).$$
In Case 2 we observe that $x_F(e_i) + x_F(e_j) = 1 + \alpha$ for some $\alpha \in [0,1)$. Thus
$$f(P) \leq \floor{\bar{x}_I(e_i) + \bar{x}_F(e_i)  + \bar{x}_I(e_j) + \bar{x}_F(e_j)} = \bar{x}_I(e_i) + \bar{x}_I(e_j)  + 1.$$
Since $\bar{x}_F(e_i) + \bar{x}_F(e_j) \geq 1$ at least one of $\bar{x}_F(e_i) ,\bar{x}_F(e_j) $ is greater than or equal to $0.5$. Assume without loss of generality $\bar{x}_F(e_i) \geq 0.5$. Then $x_(e_i) = \bar{x}_I(e_i) + 1$. So
$$f(P) \leq \bar{x}_I(e_i) + \bar{x}_I(e_j)  + 1 = x(e_i) +\bar{x}(e_j) \leq x(e_i) + x(e_j).$$
Therefore in either case $x$ satisfies the cut constraint for $P$. Therefore $x$ is feasible for $(\overline{IP})$.
The algorithm $\cA$ is a $2$-approximation algorithm for $(\overline{IP})$.
As we have argued above, $(\overline{LP})$ can be solved in polynomial time, so step $1$ can be done in polynomial time. Step $3$ can be done in $O(|E|)$ time. Hence $\cA$ terminates in polynomial time. From lemma \ref{lemma:feasible-rounding} we see that $x$ is feasible for $(\overline{IP})$. It remains to check the approximation factor.
Let $e \in E$. If $\bar{x}(e) < 0.5$ then we have immediately $x(e) \leq \bar{x}(e)$ and so trivially we have $x(e) \leq 2\bar{x}(e)$. It remains to consider the pertinent case when $\bar{x}(e) \geq 0.5$. Therein we have
\frac{x(e)}{\bar{x}(e)} &= \frac{\bar{x}_I(e) + 1}{\bar{x}_I(e) + \bar{x}_F(e)} \\
&\leq  \frac{\bar{x}_I(e) + 1}{\bar{x}_I(e) +0.5} \\
&= \frac{2\bar{x}_I(e) + 2}{2\bar{x}_I(e) + 1}.
Since $\frac{2\bar{x}_I(e) + 2}{2\bar{x}_I(e) + 1}$ is a non-increasing function, it is maximized when $\bar{x}_I(e) = 0$. Hence
$$\frac{2\bar{x}_I(e) + 2}{2\bar{x}_I(e) + 1} \leq \frac{2\cdot 0 + 2}{2\cdot 0 + 1} = 2.$$
Thus $x(e) \leq 2\bar{x}(e)$ as desired. Therefore
$$\sum_{e\in E}c(e) x(e) \leq \sum_{e \in E}c(e)2\bar{x}(e) = 2\sum_{e\in E}\bar{x}(e) \leq 2\cdot OPT,$$
where $OPT$ is the cost of an optimal solution to $(\overline{IP})$ (the last inequality follows immediately from $\bar{x}$ being optimal for $(\overline{LP})$. So $\cA$ is a $2$-approximation algorithm.
%3b
\subsection*{b}
